---
title: 'UE20CS312 - Data Analytics Worksheet 2b : Multiple Linear Regression'
author: "GAURAV MAHAJAN"
date: '2022-09-15'
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
###Importing libraries and uploading the dataset
```{r}
library(tidyverse)
library(corrplot)
library(olsrr)
df <- read_csv('spotify.csv')
head(df)
```

###Problem-1 (0.5 Points)
Check for missing values in the dataset and normalize the dataset.
```{r}
#checking for missing values
sum(is.na(df))
#Normalisation
min_max_norm <- function(x) {
  (x - min(x)) / (max(x) - min(x))
}
df_norm <- as.data.frame(lapply(df, min_max_norm))
```
This implies there is no missing data in the dataset
```{r}
#for scaling : 
#for z score scaling to be done centering is done
df<-as.data.frame(scale(df))
```

###Problem-2 (2 Points)
Fit a linear model to predict the energy rating using all other attributes.Get the summary of the model and
explain the results in detail.[Hint : Use the lm() function. Click here To get the documentation of the same.]
```{r}
#For all the attributes fitting a linear model
full_model<-lm(energy~.,data = df)
summary(full_model)
```
The asterisks tell us the significance .If alpha is 0.05 :: we select : danceability , loudness , acousticness , instrumentalness , valence according to this model.Betas are not all zero seeing he F statistic

###Problem-3 (2 points)
With the help of a correlogram and scatter plots, choose the features you think are important and model an
MLR. Justify your choice and explain the new findings.
```{r}
df_cor <- cor(df_norm)
corrplot(df_cor, order = "hclust", method = "number")
plot(df_norm)
#energy has significant positive correlation with loudness, valence and tempo
#energy has significant negative correlation with acousticness, instrumentalness
#these variables are significant in the model

model_cor <- lm(energy~loudness+acousticness+valence+tempo+instrumentalness, data=df_norm)
summary(model_cor)
```


###Problem-4 (1 Point)
Conduct a partial F-test to determine if the attributes not chosen by you in Problem-3 are significant to
predict the energy.What are the null and alternate hypotheses? [ Hint : Use the anova function between
models created in Problem-2 and Problem-3 ]
```{r}
anova(model_cor,full_model)
```


###Problem-5 (1.5 Points)
AIC - Akaike Information Criterion is used to compare different models and determine the best fit for the
data. The best-fit model according to AIC is the one that explains greatest amount of variation using the
fewest number of attributes. Check this resource to learn more about AIC.
Build a model based on AIC using Stepwise AIC regression.Elucidate your observations from the new model.
( Hint : Use an appropriate function in olsrr package.)
```{r}
stepwise_model<-lm(energy ~ loudness + acousticness + danceability + valence + instrumentalness + mode + key , data=df)
summary(stepwise_model)
```
###Problem-6 (1 Point)
Plot the residuals of the models built till now and comment on it satisfying the assumptions of MLR.
```{r}
plot(full_model$residuals , pch = 16, col="red")
abline(h=0,lty=2)
ols_plot_resid_hist(full_model)
```
```{r}
plot(model_cor$residuals , pch = 16, col="red")
abline(h=0,lty=2)
ols_plot_resid_hist(model_cor)
```

Problem-7 (2 Points)
For the model built in Problem-2 , determine the presence of multicollinearity using VIF. Determine if
there are outliers in the data using Cook’s Distance. If you find any , remove the outliers and fit the model
for Problem-2 and see if the fit improves. [ Hint : All the relevant functions can be found in olsrr package.
An observation can be termed as an outlier if it has a Cook’s distance of more than 4/n where n is the
number of records.]
```{r}
ols_vif_tol(full_model)
cooks <- ols_plot_cooksd_chart(full_model)
```
```{r}
new_df<-df[-c(30,35,49,84,114,120,127,144,153,159,171,172,187,191),]
new_full_model<-lm(energy~.,data=new_df)
summary(new_full_model)
```

