## Loading the Dataset
```{r dataload, message=FALSE, warning=FALSE, results=TRUE}
library(tidyverse)
spotify_df <- read_csv('spotify.csv')
```

## Problem-1    (0.5 Points)

Check for missing values in the dataset and normalize the dataset. 
(centralised the means or use z score test)

```{r}
library(tidyverse)
library(corrplot)
library(olsrr)

colSums(is.na(spotify_df))
```
#no missing data in the dataset
```{r}
#for scaling : 
#for z score scaling to be done centering is done
spotify_df<-as.data.frame(scale(spotify_df))
```

## Problem-2    (2 Points)

Fit a linear model to predict the *energy* rating using *all* other
attributes.Get the summary of the model and explain the results in detail.[*Hint*
: Use the lm() function. [Click
here](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/lm)
To get the documentation of the same.]

```{r}
full_model<-lm(energy~.,data = spotify_df)  #for all 
summary(full_model)
```
#stars tells us the significance . if alpha is 0.05 :: we select : danceability , loudness , acousticness , instrumentalness , valence according to this model

#betas are not all zero seeing he F statistic


## Problem-3    (2 points)

With the help of a correlogram and scatter plots, choose the features you
think are important and model an MLR. Justify your choice and explain the new findings.
```{r}
correlation<- cor(spotify_df)
corrplot(correlation,method='number')
```
```{r}
plot(x=spotify_df$loudness , y=spotify_df$energy , xlab="loudness" , ylab="energy" , main = "Energy vs Loudness")

```
```{r}
reduced_model<-lm(formula = energy ~ loudness , data = spotify_df)
summary(reduced_model)
```

## Problem-4    (1 Point)

Conduct a partial F-test to determine if the attributes not chosen by
you in *Problem-3* are significant to predict the energy.What are the
null and alternate hypotheses? [ *Hint* : Use the anova function between
models created in *Problem-2* and *Problem-3*]

```{r}
anova(reduced_model,full_model)
```
#reject the null hypothesis therefore there were a few significant features that have not been considered

## Problem-5    (1.5 Points)

AIC - Akaike Information Criterion is used to compare different models
and determine the best fit for the data. The best-fit model according to
AIC is the one that explains greatest amount of variation using the
fewest number of attributes. Check
[this](https://www.scribbr.com/statistics/akaike-information-criterion/)
resource to learn more about AIC.

Build a model based on AIC using Stepwise AIC regression.Elucidate your observations from the new model. ( *Hint* : Use
an appropriate function in
[olsrr](https://www.rdocumentation.org/packages/olsrr/versions/0.5.3)
package.)

##AIC --> encourges a better performance model
##BIC --> penalizes the complex model

##therefore : AIC and BIC should be as less as possible for it to be a proper model


```{r}
stepwise_model<-lm(energy ~ loudness + acousticness + danceability + valence + instrumentalness + mode + key , data=spotify_df)
summary(stepwise_model)
```
```{r}
plot(full_model$residuals , pch = 16, col="red")
abline(h=0,lty=2)
ols_plot_resid_hist(full_model)

```

```{r}
plot(reduced_model$residuals , pch = 16, col="red")
abline(h=0,lty=2)
ols_plot_resid_hist(reduced_model)
```
```{r}
plot(stepwise_model$residuals , pch = 16, col="red")
abline(h=0,lty=2)
ols_plot_resid_hist(stepwise_model)
```

## Problem-7    (2 Points)
For the model built in **_Problem-2_** , determine the presence of multicollinearity using VIF. Determine if there are outliers in the data using [Cook's Distance](https://www.statisticshowto.com/cooks-distance/). If you find any , remove the outliers and fit the model for _Problem-2_ and see if the fit improves. [ _Hint_ : All the relevant functions can be found in _olsrr_ package. An observation can be termed as an outlier if it has a Cook's distance of more than 4/n where n is the number of records.] 

```{r}
ols_vif_tol(full_model)
```
##everything is below 4 so nothing is violating 

```{r}
cooks <- ols_plot_cooksd_chart(full_model)
```
```{r}
new_df<-spotify_df[-c(30,35,49,84,114,120,127,144,153,159,171,172,187,191),]
new_full_model<-lm(energy~.,data=new_df)
summary(new_full_model)
```

