---
title: "UE20CS312 - Data Analytics - Worksheet 3a - Basic Forecasting Techniques"
author: "GAURAV MAHAJAN"
date: '2022-09-28'
output:
  word_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(forecast)
df <- read.csv('sales.csv')
head(df)
```
```{r}
sales <- df$Sales
head(sales)
```
```{r}
sales_ts <- ts(sales, frequency = 52, start=c(2010, 2, 5))
sales_ts
```
```{r}
plot.ts(sales_ts)
```

###Problem 1 (1 Point)
Decompose the Sales column into trend, seasonal and random components. Plot these components as well
(Hint: Look at the decompose function).
```{r}
sales_ts_component<-decompose(sales_ts)
plot(sales_ts_component)
```

###Problem 2 (3 Points)
• Perform forecasts using Single, Double and Triple Exponential Smoothing.
• Plot the forecasts of all three forecasts (different colours) against the true values. (Hint: use lines)
• Use only one function needed for all 3 forecasts, only changing parameters to get each of the 3
models (Hint: Think about alternate names)
```{r}
Single_exponential_smoothing <- HoltWinters(sales_ts, alpha=0.2, beta=0, gamma=0)

plot(sales_ts)
lines(Single_exponential_smoothing$fitted[,1], lty=2, col="blue")
```
```{r}

Single_exponential_smoothing.pred <- predict(Single_exponential_smoothing, 56 , prediction.interval = TRUE, level=0.95)
#Visually evaluate the prediction
plot(sales_ts , xlim=c(2010,2013.5))
lines(Single_exponential_smoothing$fitted[,1], lty=2, col="blue")
lines(Single_exponential_smoothing.pred[,1], col="red")
lines(Single_exponential_smoothing.pred[,2], lty=2, col="orange")
lines(Single_exponential_smoothing.pred[,3], lty=2, col="orange")
```

```{r}

Double_exponential_smoothing <- HoltWinters(sales_ts, alpha=0.2, beta=0.1, gamma=0)

plot(sales_ts)
lines(Double_exponential_smoothing$fitted[,1], lty=2, col="red")
```

```{r}
Double_exponential_smoothing.pred <- predict(Double_exponential_smoothing, 56 , prediction.interval = TRUE, level=0.95)

plot(sales_ts , xlim=c(2010,2013.5))
lines(Double_exponential_smoothing$fitted[,1], lty=2, col="blue")
lines(Double_exponential_smoothing.pred[,1], col="red")
lines(Double_exponential_smoothing.pred[,2], lty=2, col="orange")
lines(Double_exponential_smoothing.pred[,3], lty=2, col="orange")
```


```{r}

Triple_exponential_smoothing <- HoltWinters(sales_ts, alpha=0.2, beta=0.1, gamma=0.1)

plot(sales_ts)
lines(Triple_exponential_smoothing$fitted[,1], lty=2, col="orange")
```

```{r}
Triple_exponential_smoothing_pred <- predict(Triple_exponential_smoothing, 56 , prediction.interval = TRUE, level=0.95)
plot(sales_ts , xlim=c(2010,2013.5))
lines(Triple_exponential_smoothing$fitted[,1], lty=2, col="blue")
lines(Triple_exponential_smoothing_pred[,1], col="red")
lines(Triple_exponential_smoothing_pred[,2], lty=2, col="orange")
lines(Triple_exponential_smoothing_pred[,3], lty=2, col="orange")
```

###Problem 3 (2 Points)
• Forecast Sales values by Regression using all other columns. Print summary of regression model.
• Plot the predicted values against original as well. (Hint: Regression model predictions will not be a
Time Series, so need to get both to common index/x-axis)
• (Hint: Will not work unless one column is dropped/transformed before including it in the regression.
Use the lm function to get linear model)
Note: This is Multiple Linear Regression, that is, using all the columns for regression
```{r}
lm_pred <- lm(formula=Sales ~ .-X-Date,data=df)
summary(lm_pred)
#Plot the predicted values against original
library(ggplot2)
ggplot(df, aes(x=predict(lm_pred), y= Sales)) +
geom_point() +
geom_abline(intercept=0, slope=1) +
labs(x='Predicted Values', y='Actual Values', title='Predicted vs. Actual Values')
```


###Problem 4 (2 Points)
Plot the Laptop_Demand column as a time series. Identify the forecasting required for this type of Time-series,
and forecast the values for all 143 weeks (Hint: Look at functions in the forecast package)
```{r}
library(forecast)

laptop <- df$Laptop_Demand
laptop_ts <- ts(laptop, frequency = 12)


plot(laptop_ts)

### ARIMA is used to forecast the above timeseries as we are predicting futur values using the past.
mymodel <- auto.arima(laptop_ts)

forecast_future <- forecast(mymodel, level=c(95), h=143)
forecast_future
```
###Problem 5 (2 Points)
Evaluate the accuracy of all 3 Exponential Smoothing models (from Problem 2) and Regression models
using the MAPE and RMSE metrics. Comment on which is the best Exponential Smoothing method, and
if Regression is better than Exponential Smoothing? Provide a reasoning for why the best model is better
suited for the Sales data (Bonus Point: reasoning for why the 2 other models perform similarly)
```{r}
library("Metrics")
#For regression Model
sprintf("MAPE value for regression model is: %f",mape(sales_ts,lm_pred$fitted))
sprintf("RMSE value for Regression model is: %f",rmse(sales_ts,lm_pred$fitted))

#For Single Exponential Smoothing Model
sprintf("MAPE value for Single Exponential smoothing model is: %f",mape(sales_ts,Single_exponential_smoothing$fitted))
sprintf("RMSE value for Single Exponential smoothing model is: %f",rmse(sales_ts,Single_exponential_smoothing$fitted))

#For Double Exponential Smoothing Model
sprintf("MAPE value for Double Exponential smoothing model is: %f",mape(sales_ts,Double_exponential_smoothing$fitted))
sprintf("RMSE value for Double Exponential smoothing model is: %f",rmse(sales_ts,Double_exponential_smoothing$fitted))

#For Triple Exponential Smoothing Model
sprintf("MAPE value for Triple Exponential smoothing model is: %f",mape(sales_ts,Triple_exponential_smoothing$fitted))
sprintf("RMSE value for Triple Exponential smoothing model is: %f",rmse(sales_ts,Triple_exponential_smoothing$fitted))
```

