---
title: "UE20CS312 - Data Analytics - Worksheet 3b - AR and MA models"
author: "GAURAV MAHAJAN"
date: '2022-10-16'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
###Problem 0 (0.5 point)
Set the index of DataFrame to the Date column to make it a time series
```{r}

```
##Problem 1 (1.5 point)

Plot the time-series. Analyze the stationarity from the time-series. Provide reasoning for stationarity/non-
stationarity based on visual inspection of time-series plot (0.5 point)
```{r}

```
Plot the ACF plot of the Time series (upto 50 lags). Analyze the stationarity from ACF plot and provide
reasoning (Hint: look at functions in statsmodels package) (1 Point)
```{r}

```
###Problem 2 (2 point)
Run Augmented Dickey Fuller Test. Analyze whether the time-series is stationary, based on ADF results
(1 point)
Hint: Use the print_adf_results function below to print the results of the ADF function cleanly after
obtaining results from the library function. Pass the results from library function to print_adf_results
function
```{r}

```
Problem 3 (1 point)
Plot both ACF and PACF plot. From these select optimal parameters for the ARIMA(p,q) model
Hint: Negative values that are significantly outside the Confidence interval are considered significant too.
Hint: p+q = 3
```{r}

```

```{r}

```

###Problem 4 (2 point)
Write a function to forecast values using only AR(p) model (2 Points)

Only use pandas functions and Linear Regression from sklearn . LR documentation (https://scikit-
learn.org/stable/modules/generated/sklearn.linear_model.LinearRegression.html)

Hint: Create p new columns in a new DataFrame that is a copy of transformed_df
Each new column has lagged value of Price. Price_t-i (From Price_t-1 upto Price_t-p )
Look at the shift function in pandas to create these new columns Link
(https://pandas.pydata.org/docs/reference/api/pandas.DataFrame.shift.html)
```{r}

```
###Problem 5 (1 Point)
Phew! Just handling AR(2) manually required us to difference, apply regression, undifference. Let's make all
of this much easier with a simple library function
Use the ARIMA function using parameters picked to forecast values (1 point)
```{r}

```
###Problem 6 (1 point)
Evaluate the ARIMA model using Ljung Box test. Based on p-value infer if the Model shows lack of fit
